---
title: "2. Series exploration"
author: "Andrei Pazniak"
date: "2022-09-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, echo=FALSE, include=FALSE}
library(data.table);library(ggplot2)
source("..\\common\\QuotesTransform.R")

data <- fread("cleanedSeries.csv")
targetName <- "XAUUSD"
```


We want to predict future y with current regressor x. Simulating basic examples and using cross correlation function.
```{r}
set.seed(111)
x1<-frollsum(rnorm(100), 10)
y1 <- shift(x1, 10, type="lag")^3  # y1[i+10] = x1[i]^3 - we can predict the future with x

ccf(y1, x1, lag.max = 75, na.action = na.pass)
```


# Analysing series for `r `targetName`

Cross correlation will be the first exploration. Just some notes for analysing:
- If correlation high/low for lag > 0, it means predictor forecast `target` for "Lag" days. `targetName` will change to the same side as predictor if corr is high (or opposite side if predictor is negative low)
- If correlation high/low for lag < 0, consider to exchange target and regressor variable.
```{r example}
scale_max_value <- data[series_id==targetName][.N, value]
data[type=="Price/USD", y := Scale_Max(m2adjusted_value, scale_max_value), by=series_id]
data[type=="USA Economic/Percent", y := Scale_Max(value, scale_max_value), by=series_id]

for(curr_id in unique(data$series_id))
  ccf(data[series_id==targetName, y], data[series_id==curr_id, y], lag.max=6000, na.action=na.pass, main=curr_id)
```




correlation
```{r}
#d<-rbind(sp500[,.(date, series_id, value=growthIndex)], fed_ii)
#d[series_id=="^GSPC", value:=value*10]
#cor( sp500[series_id=="^GSPC", m2adjusted_value], econ_parameters[series_id=="FEDFUNDS", #value], #use="complete.obs")
#cor( sp500[series_id=="^GSPC", m2adjusted_value], econ_parameters[series_id=="UNRATE", #value], #use="complete.obs")
#cor( econ_parameters[series_id=="FEDFUNDS", value], econ_parameters[series_id=="UNRATE", value], #use="complete.obs")
#ccf(econ_parameters[series_id=="FEDFUNDS", value], econ_parameters[series_id=="UNRATE", value])
#ccf(econ_parameters[series_id=="FEDFUNDS", value], econ_parameters[series_id=="T10Y2Y", value], #lag.max=6000, na.action=na.pass)


#scale to latest sp500 value

#acf(econ_parameters[series_id=="UNRATE", value], lag.max = 20000)
```

Linear regresion
```{r lr}
#lm( sp500[series_id=="XAUUSD", m2adjusted_value]~ econ_parameters[series_id=="FEDFUNDS", value] + #econ_parameters[series_id=="UNRATE", value] + econ_parameters[series_id=="T10Y2Y", value])

```

