---
title: "5.SPXAnalysis"
author: "Andrei Pazniak"
format: html
editor: visual
---

```{r init, echo=FALSE, include=FALSE}
library(data.table);library(ggplot2);library(plotly)
source("..\\common\\QuotesTransform.R")

data <- fread("cleanedSeries.csv")

targetName <- "^GSPC"

```

### TLT Price normal and adjusted

```{r}
p <- ggplot(data[series_id==targetName])  + geom_line(aes(x=date, y=adjusted_value_m2, color='Adjusted_M2 SP500')) +geom_line(aes(x=date, y=value, color="SP500")) + ggtitle("SP500 Price normal and adjusted") + ylab(targetName) + xlab("date")
ggplotly(p)
```

### Correlation Treasuries and FEDFUNDS

```{r}
data[series_id==targetName, y := growthIndex]
data[series_id=="FEDFUNDS", y := value/10]
d <- data[series_id==targetName | series_id == "FEDFUNDS", .(series_id, date, y)]

d2 <- d[series_id == "FEDFUNDS"]
d2[, series_id:="FEDFUNDS_MA200"]
d2[, y:= frollmean(y, 200, na.rm = T)]

d <- rbind(d, d2)


p <- ggplot(d, aes(x=date, y=y, color=series_id)) + geom_line() + ggtitle("SPX and FEDFUNDS") + ylab("growth Index/yield*0.1") + xlab("date")

ggplotly(p)

```

### FedFund trends

There are 3 states of FedFunds:

1.  Trend Up (If diffM200 for fedfunds is greater than threshold)

2.  Trend Down (If diffM200 for fedfunds is less than threshold)

3.  Sideway (If diffM200 for abs(fedfunds) is less than threshold)

To find threshold we make an assumption based on previous chart that 10% of time market in sideways and 90% in trend.

```{r}
fed_Sideways_Threshold <- quantile(d[, abs(valueMVA-value_fed)], probs=c(0.25), na.rm=TRUE)
print(paste("Threshold level for classification sideway is difference between moving average 200 and real value", fed_Sideways_Threshold[[1]]))

d_trend <- data[series_id == "FEDFUNDS", .(value, date, valueMVA)]
d_trend[value - valueMVA>=fed_Sideways_Threshold[[1]], trend:="up"]
d_trend[value - valueMVA<=-fed_Sideways_Threshold[[1]], trend:="down"]
d_trend[is.na(trend), trend:="sideway"]

d_sp <- data[series_id == targetName, .(value, date )]
d <- merge(d_sp, d_trend[, .(date, value_fed=value, trend)], by="date", all.x=T, all.y = F)


d_wide <- dcast(d, date  ~ trend, value.var = "value_fed")
d_long <- melt(d_wide, id.vars = c("date"), variable.name = "trend", value.name = "value")

p <- ggplot(d_long) + geom_line(aes(x=date, y=value, color=trend)) + geom_line(aes(x=date, y=value), alpha = 0.1) + ggtitle("Trends in FEDFUNDS") + ylab("growth Index/yield*0.1") + xlab("date")

ggplotly(p)
```

After trend classification we can see SPX ~ FED in convienient chart


```{r}
d_wide <- dcast(d, date  ~ trend)
d_long <- melt(d_wide, id.vars = c("date"), variable.name = "trend", value.name = "value")

p <- ggplot(d_long) + geom_line(aes(x=date, y=value, color=trend)) + geom_line(aes(x=date, y=value), alpha = 0.1) + ggtitle("SPX and FEDFUNDS") + ylab("price") + xlab("date")

ggplotly(p)
```

SP 500 growthIndex  with colors

```{r}
d_sp <- data[series_id == targetName, .(growthIndex, date )]
d_growthIndex <- merge(d_sp, d_trend[, .(date, trend)], by="date", all.x=T, all.y = F)

d_wide <- dcast(d_growthIndex, date  ~ trend, value.var = "growthIndex")
d_long <- melt(d_wide, id.vars = c("date"), variable.name = "trend", value.name = "growthIndex")

p <- ggplot(d_long) + geom_line(aes(x=date, y=growthIndex, color=trend)) + geom_line(aes(x=date, y=growthIndex), alpha = 0.1) + ggtitle("SPX drawdown and FEDFUNDS") + ylab("drawdown") + xlab("date")

ggplotly(p)
```


todo refactor tp russel2000 \### TLT, Junk bond and their ratio

```{r bondyield}
bond <- data[date>'1998-01-01'][series_id=="BAMLH0A0HYM2EY" | series_id=="IRLTLT01USM156N"]
scaleFactor <- 10
d <- data.table(date = unique(bond$date), value = scaleFactor*bond[series_id=="IRLTLT01USM156N", value]/ bond[series_id=="BAMLH0A0HYM2EY", value], series_id = "tlt2junk", type="USA Economic/Percent")
d <- rbind(bond, d, fill=TRUE)
p <- ggplot(d, aes(x=date, y=value, color=series_id))  + geom_line() + geom_hline(yintercept = 2, color="blue") + geom_hline(yintercept = 5, color="blue") + ggtitle("TLT, Junk bond and their ratio") + ylab("yield/ratio") + xlab("date")
ggplotly(p)

```
